---
title: "HBO2"
author: "Philippe"
date: "26 November 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---


The aim is to analyse the effect of group composition and breeding status of the colony on the behavioural response of resident to male immigration.

```{r setup, include=FALSE}
rm(list=ls())#clean environment

library(RMySQL)
library(getPass)
library(EloRating)
library(tidyverse)
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(hms)
```

We start by establishing a connection with the database
```{r Database connection}
con <- dbConnect(MySQL(), user = 'philippev',password = getPass(),  
                 dbname = 'Moleratdatabase', host = 'Kalahariresearch.org')

#dbDisconnect(con)

#set seed to get always the same randomization
set.seed(123)
```

The first thing that one has to do is to extract the information on Colony composition on the day of observation 

```{r extract group composition info}

Membership <- con %>%
  dbGetQuery ("SELECT 
    AnimalRef,
    AnimalID,
    DATE(MemberFrom) AS MemberFrom,
    DATE(MemberTo) AS MemberTo,
    MemberShipBetween.ColonyRef,
    MemberDays,
     MemberShipBetween.Colony AS QueriedColony,
    tblColonyCodes.ColonyOrigin
FROM
    MoleratViews_Pending.MemberShipBetween
LEFT JOIN
    Moleratdatabase.tblColonyCodes ON MoleratViews_Pending.MemberShipBetween.ColonyRef = tblColonyCodes.ColonyRef
WHERE MemberShipBetween.ColonyRef <> 120 
AND MemberShipBetween.Colony <> 'Exported_Nigel'") %>% 
  mutate(MemberFrom=ymd(MemberFrom),MemberTo=ymd(MemberTo)) %>% 
  select(AnimalRef,AnimalID,MemberFrom,MemberTo,QueriedColony,ColonyOrigin)

#get sex from the database
Sex<-con %>% 
  dbGetQuery("SELECT *
FROM Moleratdatabase.tblSex
") %>% 
  select(AnimalID,Sex)

#get the weight from the database
WeightList<-con %>% 
  dbGetQuery("SELECT * FROM user_philippev.Weight_AnimalID") %>% 
  mutate(WeightDate=ymd(WeightDate)) %>% 
  select(AnimalID,WeightDate,Weight,WeightType) %>% 
  filter(!(is.na(Weight)))

#get the id characteristic from the database 
ID_Characteristic <-
  con %>% 
  dbGetQuery("SELECT * FROM user_philippev.ID_Characteristic") %>% 
  mutate(BirthDate=ymd(BirthDate),DeathDate=ymd(DeathDate),Mother_FirstLitter=ymd(Mother_FirstLitter),Father_FirstLitter=ymd(Father_FirstLitter)) %>% 
  select(AnimalID,Sex,Wildcaught,WildcaughtQueen,BirthDate,LitterRef,Mother_FirstLitter,Father_FirstLitter,DeathDate)

```

The first step is to get the focal Session 
```{r focal data}

#get the focal sessions from the database 
Focal<-con %>% 
  dbGetQuery("SELECT * FROM user_philippev.HBO2_FocalSession") %>% 
  rename(Date=ObsDate,Colony=ResidentColony) %>% 
  mutate(Treatment=as.factor(Treatment),Date=ymd(Date))

```

The second step is to get the group composition data
```{r}
source("Group_functions.R")

#the colony has already been extracted so we can go straight to the group comp 
List<-Focal %>% 
  select(Colony,Date)
names(List)

#get GroupComp
GroupComp<-get_GroupComp(List,Membership)
names(GroupComp)
nrow(GroupComp)


#add the closest Weight to all animals within the group
#we lose 3 lines, because 3 animals have no weigth thus no date of weight and thus no timediff can be computed 
Weight<-get_Weight(GroupComp %>% select(AnimalID,Date),WeightList)
names(Weight)
nrow(Weight)

#get individual information 
ID_Info<-get_IDinfo(GroupComp %>% select(AnimalID,Date),ID_Characteristic)
View(ID_Info)
nrow(ID_Info)#all matches all good 
names(ID_Info)

names(ID_Characteristic)

#Bring all info together before calculating the characteristic of the group 
#On the database, I would need to add the paternity of originally wildcaught colony and make sure the paternity of lab formed colony is exact
#what to do with age rank and individual of unknown age? They are older but they cannot be distinguished with one another. 
Group_ID_Info <-
  left_join(GroupComp,Weight) %>% 
  left_join(.,ID_Info)

View(Group_ID_Info)


#get the additional information of the group at every day of observation
#in case one may need to add rank after removing the breeders: 
Group_AllInfo<-get_GroupID_Info(Group_ID_Info) %>% 
  group_by(Colony,Date) %>% 
  mutate(Male09=sum(Sex=="M" & (Age >9|is.na(Age))),Male12=sum(Sex=="M" & (Age >12|is.na(Age))),Male15=sum(Sex=="M" & (Age >15|is.na(Age))),Male18=sum(Sex=="M" & (Age >18|is.na(Age)))) %>% 
    mutate(Female09=sum(Sex=="F" & (Age >9|is.na(Age))),Female12=sum(Sex=="F" & (Age >12|is.na(Age))),Female15=sum(Sex=="F" & (Age >15|is.na(Age))),Female18=sum(Sex=="F" & (Age >18|is.na(Age)))) %>% 
  mutate(MaleNB=sum(Sex=="M"), FemaleNB=sum(Sex=="F")) %>% 
  ungroup()

names(Group_AllInfo)

```

## Data Exploration 
Before we carry out any analyses we shall have a look at what our sample size 

the first analysis is to see how the duration of the observation varies as a function of treatment and group characteristic 

```{r}
names(Colony)

Colony<-inner_join(Focal,Group_AllInfo %>% select(Colony,Date,PupPresence,GroupSize,Male09:FemaleNB) %>% 
    distinct()) %>% 
  mutate(Treatment = factor(Treatment, levels = c( "Breeding", "Queenless","Kingless","Breederless"))) %>% 
  gather(.,Subset,Count,Male09:FemaleNB)
View(Colony)
str(Colony)

# distribution of the group size
ggplot(Colony,aes(Treatment,GroupSize))+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(width=0.1,height=0.2,size=2,col=1,alpha=0.2)+
  theme_classic()

#distribution of male number and female number 
ggplot(Colony,aes(Treatment,Count))+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(width=0.1,height=0.2,size=2,col=1,alpha=0.2)+
  facet_wrap(.~Subset, ncol = 5)+
  theme(axis.text.x= element_text(size=4))


```



```{r duration data}
#get all the duration data together. 
Duration<-inner_join(Focal,Group_AllInfo %>% select(Colony,Date,PupPresence,GroupSize,Male9:FemaleNB) %>% 
    distinct()) %>% 
mutate(Treatment = factor(Treatment, levels = c( "Breeding", "Queenless","Kingless","Breederless")))
nrow(Duration)



#first visualise the data
ggplot(Duration,aes(Treatment,Duration))+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(width=0.1,height=0.2,size=2,col=1,alpha=0.2)+
  theme_classic()

#it looks like there is a bimodal distribution . Either the colony defend or they don't. It would be nice to see what makes the difference 
ggplot(Duration,aes(Treatment,Duration))+
  geom_violin()

names(Duration)

ggplot(Duration,aes(MaleNB,Duration,col=Treatment))+
  geom_point()+
  geom_smooth(method = "lm",se=FALSE)+
  theme_classic()
 





```






#effect on mating from resident (reflects how many times the intruder could mate)
total number of mating within a colony
percentage of the female present in the group that could mate

##Effect on aggressive behaviours 

